### 2024-12-23 Manuscript files/title.tex
Title page (to be included manually)

\documentclass{article}
\usepackage{graphicx}

\begin{document}

\begin{titlepage}
    \thispagestyle{empty} 
    \centering
    \vspace*{1 cm}

    {\LARGE \textbf{What Really Matters (and to Whom): Investigating Heterogeneous Effects in High-Dimensional Conjoint Analyses}}\\[1.5 cm]

    {\Large \textbf{Thomas Reiter} \quad \textbf{Philipp Sterner}}\\[1 cm]

    {\large Statistical Consulting, Master Statistics and Data Science\\
    LMU Munich, Institute of Statistics}\\[1.5 cm]

    {\large \textbf{Supervisor:} Prof. Dr. David Rügamer}\\[1 cm]

    {\large \textbf{Project Partners:} Prof. Dr. Paul Thurner and Fabian Haggerty\\
    (Geschwister-Scholl Institute of Political Science)}\\[2 cm]

    {\large December 23, 2024}

    \vfill
\end{titlepage}
\clearpage
\pagenumbering{arabic}

\end{document}